---
title: "Report on association of systolic blood pressure and death with data from the Framingham study"
author: "Nicole Schönenberger"
date: "`r date()`"
format:
  html:
    self-contained: true
    code-link: true
    code-fold: true
    code-summary: "Code"
    toc: true
    df-print: paged
    embed-resources: true
---

## Introduction

This short report studies the  systolic blood pressure in the dataset of the Framingham study and its association with death. The Framingham data can be obtained from the RiskCommunicator package of R. Details of this package can be found in:  <https://cran.r-project.org/web/packages/riskCommunicator/index.html>.
In this analysis, only the baseline data (timepoint = 0) is used. 

## Data import, setting working directory, and loading libraries
In a first step, data is imported and only the baseline data is selected. Additionally, libraries are loaded and the working directory is set.

```{r}
#| output: false
# Working directory
setwd("~/BSAPR Assessment")

# Loading packages
library(riskCommunicator)
library(tidyverse)

# Framingham heart study data
framData <- riskCommunicator::framingham

# We keep only the baseline data
framData_base <- subset(framData, TIME == 0)

```

## Descriptive statistics

In this section, the descriptive statistics of the sample systolic blood pressure is calculated. This includes sample mean, median, quartiles, sample variance, sample standard deviation, range, interquartile range (IQR).

```{r}
# Calculating the mean
mean_bp <- mean(framData_base$SYSBP, na.rm = TRUE)
mean_bp

# Calculating the median
median_bp <- median(framData_base$SYSBP, na.rm = TRUE)
median_bp

# Calculating quartiles
quart_bp <- quantile(framData_base$SYSBP, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)
quart_bp

# Calculating sample variance
var(framData_base$SYSBP, na.rm = TRUE)

# Calculating standard deviation
sd_bp <- sd(framData_base$SYSBP, na.rm = TRUE)
sd_bp 

# Range of systolic blood pressure
range(framData_base$SYSBP, na.rm = TRUE)

# Calculating interquartile range
IQR(framData_base$SYSBP, na.rm = TRUE)

# Summary of descriptive statistics
summary(framData_base$SYSBP)
```

The `echo: false` option disables the printing of code (only output is displayed).
